### 2016-06-29 stitched.tex
ENH:  Evaluation results.

\usepackage{booktabs}
\usepackage[final]{changes}
\usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
\linespread{1.2}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\setlist{nolistsep}
\titlespacing{\section}{2pt}{*0}{*0}
\titlespacing{\subsection}{2pt}{*0}{*0}
\titlespacing{\subsubsection}{2pt}{*0}{*0}
\setlength{\parskip}{3pt}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Nick Tustison}, color=red]{nt}
\definechangesauthor[name={James Stone}, color=magenta]{js}
\definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
\definechangesauthor[name={Andy Mayer}, color=green]{am}
\definechangesauthor[name={Harvey Levin}, color=blue]{hl}
\definechangesauthor[name={Brian Taylor}, color=purple]{bt}
\definechangesauthor[name={David Tate}, color=orange]{dt}

@@

@@

@@

In Figure 8 are the segmentation comparisons derived from manual
segmentations of the same data. Despite the large variability
characteristic with manual labelings in related
fields\textsuperscript{37,39,40}, such labelings are characteristic of
current clinical practices and the methodology proposed herein is
readily adapted to refinements in training data.

On the left of Figure 8 are the improvement in Dice
values\textsuperscript{41}, over all white matter hyperintensities when
comparing the segmentations between the two stages where the sum is
taken over all individually labeled manual, \(T_r\), and automated,
\(S_r\), lesions and \(\cap\) represents the intersection between the
manual/automated lesion pair. Performing the second round of supervised
learning improves these Dice values. One can also note from the right
side of Figure 7 that the total lesion load volume illustrates a few
subjects that are severe outliers in terms of the number of false
positives. The second round helps to correct this issue.

@@

Two major meta-analyses of WMHs have been published covering the periods
prior to 2010\textsuperscript{42} and after\textsuperscript{43}. The
earlier meta-analysis covered 53 longitudinal studies that included
samples of high-risk populations, i.e., patients selected for a specific
disease or condition such as hypertension, whereas other studies
recruited samples of the general population. Longitudinal studies of
samples representative of the general population are more relevant to
the focus of the present paper. Debette \& Markus\textsuperscript{42}
found that the presence of WMHs was related to subsequent cognitive
decline, a higher risk of developing dementia, stroke, and of mortality.
Lesion volume at baseline was also predictive of cognitive decline.
Limitations of this meta-analysis include heterogeneity in the method of
measuring WMHs; some studies used automated volumetric measurement,
whereas others used a visual rating scale. The studies analyzed by
Debette \& Markus were limited to the occurrence of one of the
aforementioned conditions which they analyzed by hazard ratios.

The more recent meta-analysis by Kloppenborg et al.\textsuperscript{43}
of 23 cross-sectional studies reporting MRI and concurrent
neuropsychological results in patients with heterogeneous diagnoses but
without previously diagnosed cognitive impairment, found that WMHs were
associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
to -0.08) after controlling for age. These studies also differed in the
metric used to measure the WMHs, including volume, \% of total
intracranial volume, and a visual rating score. The effect size for the
association with cognitive deficit in these cross-sectional studies did
not differ significantly across various cognitive domains or the method
of measuring lesion volume. Among eight longitudinal studies analyzed by
Kloppenborg et al that included a follow-up MRI and also controlled for
age, the effect size for the association of progression in WMHs and
cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
association was stronger for attention and executive function than for
memory and processing speed. Although baseline WMHs were predictive of
cognitive deficit at follow-up in the seven studies which did not repeat
MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
in the longitudinal studies that calculated progression in WMHs. In
summary, progression of WMHs seen on repeat MRI has a stronger relation
to cognitive deficit than concurrent imaging findings. These
meta-analyses support the rationale for repeating an MRI in patients
younger than 50 years whose initial scan shows WMHs.

Despite the above-described associations between WMHs, cognitive
decline, increased risk of developing dementia, and mortality, these
lesions receive little attention in current clinical workflows. When
reported in a standard neuroradiologist interpretation, they are
typically handled as incidental findings and are assigned little
clinical significance. This likely reflects the impracticality of
performing a detailed assessment of number, volume, and distribution
within a qualitative neuroradiologist interpretation as well as the lack
of correlative information on how the presence and distribution of these
lesions may inform a diagnosis and prognosis in the appropriate clinical
setting. To date, automated or semi-automated tools for the detection of
WMHs have lacked the specificity and efficiency for the mining of
large-scale datasets to generate highly granular data on whether these
lesions possess any true diagnostic or prognostic value in the setting
of a specific disease process. The present communication describes a
supervised statistical learning tool that is appropriate for the
application to such large-scale datasets.

The currently described tool is just one example of how ``supervised
learning'' algorithms might be applied to aid in the diagnosis of TBI
and other disease processes through the specific identification of
features predictive of a given disease state. It is an important
demonstration of the potential power of these analytical approaches in
the rapid but comprehensive mining of information from neuroimaging
examinations. Supervised learning algorithms are presently employed
across a wide variety of settings for the rapid identification of
predictive imaging features\textsuperscript{44--47}. Automobile
manufacturers utilize these types of approaches to equip self-driving
vehicles to recognize and respond to unique external surroundings
through the identification of visual information sufficiently similar to
previously assimilated training data\textsuperscript{48,49}. Similarly,
in the context of the neuroimaging assessments, deep learning approaches
may allow for the rapid identification of information predictive of
disease state in an individual patient. These approaches have been
applied to the segmentation of macroscopically visible
structures\textsuperscript{44--47}. Additionally, these approaches might
be applied to the interrogation of imaging data in the individual
patient with a primary quantitative output metrics to include sequences
such as diffusion tensor imaging (DTI) and its variants, functional
connectivity, perfusion weighted imaging, and cortical thickness
assessments. At present, these advanced neuroimaging sequences are
confined to cohort-based research studies due to the lack of available
analytical tools to assess the information in the setting of the
individual patient\textsuperscript{50}. Application of deep learning
approaches in the context of data with primary quantitative outputs will
require large scale normative and disease specific databases. Building
these large scale imaging libraries is resource intensive and requires a
multi-center approach with harmonized scanners between sites and
correlative non-imaging clinical data. Large scale TBI data is becoming
increasingly available through activities such as the Chronic Effects of
Neurotrauma Consortium (CENC), Transforming Research and Clinical
Knowledge in TBI (TRACK-TBI), Collaborative European Neurotrauma
Effectiveness Research in TBI (CENTER-TBI), Department of Defense
Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and other data
being consolidated through FITBIR. In concert with any available high
quality normative neuroimaging data, deep learning algorithms may be
well positioned to help transform how neuroimaging is interpreted for
the clinical management of patients with this disease process.

@@

\hypertarget{ref-Grimaud:1996aa}{}
39. Grimaud J, Lai M, Thorpe J, Adeleine P, Wang L, Barker GJ, Plummer
DL, Tofts PS, McDonald WI, Miller DH. Quantification of mRI lesion load
in multiple sclerosis: A comparison of three computer-assisted
techniques. Magn Reson Imaging. 1996;14(5):495--505.

\hypertarget{ref-styner2008}{}
40. Styner M, Lee J, Chin B, Chin M, Commowick O, Tran H, Markovic-Plese
S, Jewells V, Warfield S, editors. Special Issue on 2008 MICCAI Workshop
- MS Lesion Segmentation. MIDAS J; 2008.

\hypertarget{ref-tustison2009}{}
41. Tustison NJ, Gee JC. Introducing Dice, Jaccard, and other label
overlap measures to ITK. Insight Journal. 2009.

\hypertarget{ref-Debette:2010aa}{}
42. Debette S, Markus HS. The clinical importance of white matter
hyperintensities on brain magnetic resonance imaging: Systematic review
and meta-analysis. BMJ. 2010;341:c3666.

\hypertarget{ref-Kloppenborg:2014aa}{}
43. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
Presence and progression of white matter hyperintensities and cognition:
A meta-analysis. Neurology. 2014;82(23):2127--38.

\hypertarget{ref-Plis:2014aa}{}
44. Plis SM, Hjelm DR, Salakhutdinov R, Allen EA, Bockholt HJ, Long JD,
Johnson HJ, Paulsen JS, Turner JA, Calhoun VD. Deep learning for
neuroimaging: A validation study. Front Neurosci. 2014;8:229.

\hypertarget{ref-Suk:2015aa}{}
45. Suk H-I, Lee S-W, Shen D, Alzheimer's Disease Neuroimaging
Initiative. Deep sparse multi-task learning for feature selection in
alzheimer's disease diagnosis. Brain Struct Funct. 2015 May.

\hypertarget{ref-Li:2014aa}{}
46. Li R, Zhang W, Suk H-I, Wang L, Li J, Shen D, Ji S. Deep learning
based imaging data completion for improved brain disease diagnosis. Med
Image Comput Comput Assist Interv. 2014;17(Pt 3):305--12.

\hypertarget{ref-Liu:2015aa}{}
47. Liu S, Liu S, Cai W, Che H, Pujol S, Kikinis R, Feng D, Fulham MJ,
ADNI. Multimodal neuroimaging feature learning for multiclass diagnosis
of alzheimer's disease. IEEE Trans Biomed Eng. 2015;62(4):1132--40.

\hypertarget{ref-HadsellSBESKML09}{}
48. Hadsell R, Sermanet P, Ben J, Erkan A, Scoffier M, Kavukcuoglu K,
Muller U, LeCun Y. Learning long-range vision for autonomous off-road
driving. J. Field Robotics. 2009;26(2):120--144.

\hypertarget{ref-FarabetCNL12}{}
49. Farabet C, Couprie C, Najman L, LeCun Y. Scene parsing with
multiscale feature learning, purity trees, and optimal covers. In:
Proceedings of the 29th international conference on machine learning,
ICML 2012, edinburgh, scotland, uK, june 26 - july 1, 2012. icml.cc /
Omnipress; 2012.

\hypertarget{ref-Mayer:2014aa}{}
50. Mayer AR, Bedrick EJ, Ling JM, Toulouse T, Dodd A. Methods for
identifying subject-specific abnormalities in neuroimaging data. Hum
Brain Mapp. 2014;35(11):5457--70.

>>>>

\usepackage{booktabs}
\usepackage[final]{changes}
\usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
\linespread{2}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\setlist{nolistsep}
\titlespacing{\section}{2pt}{*0}{*0}
\titlespacing{\subsection}{2pt}{*0}{*0}
\titlespacing{\subsubsection}{2pt}{*0}{*0}
\setlength{\parskip}{3pt}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Nick Tustison}, color=red]{nt}
\definechangesauthor[name={James Stone}, color=magenta]{js}
\definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
\definechangesauthor[name={Andy Mayer}, color=green]{am}
\definechangesauthor[name={Harvey Levin}, color=blue]{hl}
\definechangesauthor[name={Brian Taylor}, color=purple]{bt}
\definechangesauthor[name={David Tate}, color=orange]{dt}

@@

\textcolor{blue}{Several measures have been employed in the literature for evaluating
automated white matter lesion segmentation involving such quantities as lesion load,
voxel-based overlap measures (such as the Dice similarity coefficient), and lesion-based measures.}\textsuperscript{37}\textcolor{blue}{.  For this work, due to the relatively small-size distribution}\footnote{\textcolor{blue}{For comparison, the training data set of
  the MS Lesion Segmentation challenge
  associated with the international MICCAI 2008 conference has a mean lesion load of
  204 ($\pm$ 752) mm$^3$ per lesion and the resolution is almost twice what is used
  in this study (i.e., 0.5 $\times$ 0.5 $\times$ 0.5).}}
\textcolor{blue}{of the lesion load in our data cohort (see Table 1), we used
four lesion-based measures:  sensitivity, positive predictive value (i.e., precision), $F_1$ score, and relative volume difference.  The first three quantities are based on the number of false positives ($TN$, false negatives ($FN$), and true positives ($TP$) in terms of identified lesions.  It should be noted that the number of true negatives is not readily
incorporated into measures of accuracy as the quantity of ``true negatives'' (i.e., correctly identified normal brain tissue) would severely skew accuracy assessments. The $F_1$ score is an assessment of accuracy which takes into account both the sensitivity,}

\[ sensitivity =  \frac{ TP }{TP + FN}, \]

\textcolor{blue}{and the positive predictive value (PPV),}

\[ PPV =  \frac{ TP }{TP + FP}, \]

\textcolor{blue}{such that $F_1$ is given by}

\[ F_1 =  \frac{ 2 \cdot TP }{ 2 \cdot TP + FP + FN}. \]

\textcolor{blue}{The relative volume difference is calculated for each of the true positive lesions using
the manual and predicted lesion volumes:}

\[ Relative\,\,volume\,\,difference = \frac{V_{manual} - V_{predicted}}{V_{manual}}. \]

\textcolor{blue}{To illustrate how the performance of our framework varies with lesion size, we calculated
the above measures based on evenly split quantiles of the manual estimates of lesion volumes into 3 groups.  These three size ranges (in terms of the number of voxels) were:  $[3-12)$, $[12-28)$, and $[28-551]$.}

@@

\textcolor{blue}{In Figure 8 we provide the segmentation evaluations derived from the leave-one-out evaluation
of the previously described TBI data.  These performance measures include sensitivity,
positive predictive value, $F_1$ score, and relative volume difference.  The three lesion
size ranges over which these measures are computed to illustrate the variation
in performance with lesion size.  Smaller lesions ($< 12$ voxels) are more difficult
to identify which is why the sensitivity is widely varied over the cohort vs. the largest
set of voxels ($> 28$ voxels).   The first three measures are based on the identification of
entire lesions.  The relative volume difference provides a direct assessment of how accurate
the volumetric estimate is when comparing the manually identified lesions versus the
automatically predicted lesions.}

@@

@@

Two major meta-analyses of WMHs have been published covering the periods
prior to\textsuperscript{39} and after 2010\textsuperscript{40}. Debette
\& Markus\textsuperscript{39} found that the presence of WMHs was
related to subsequent cognitive decline, a higher risk of developing
dementia, stroke, and of mortality. Lesion volume at baseline was also
predictive of cognitive decline. Kloppenborg et al.\textsuperscript{40}
of 23 cross-sectional studies reporting MRI and concurrent
neuropsychological results in patients with heterogeneous diagnoses but
without previously diagnosed cognitive impairment, found that WMHs were
associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
to -0.08) after controlling for age.

Despite the potential clinical significance of WMHs these lesions
receive little attention in current clinical workflows. When reported in
a standard neuroradiologist interpretation, they are typically handled
as incidental findings and are assigned little clinical significance.
This likely reflects the impracticality of performing a detailed
assessment of number, volume, and distribution within a qualitative
neuroradiologist interpretation as well as the lack of correlative
information on how the presence and distribution of these lesions may
inform a diagnosis and prognosis in the appropriate clinical setting. To
date, automated or semi-automated tools for the detection of WMHs have
lacked the specificity and efficiency for the mining of large-scale
datasets to generate highly granular data on whether these lesions
possess any true diagnostic or prognostic value in the setting of a
specific disease process. The present communication describes a
supervised statistical learning tool that is appropriate for the
application to such large-scale datasets.

@@

\hypertarget{ref-Debette:2010aa}{}
39. Debette S, Markus HS. The clinical importance of white matter
hyperintensities on brain magnetic resonance imaging: Systematic review
and meta-analysis. BMJ. 2010;341:c3666.

\hypertarget{ref-Kloppenborg:2014aa}{}
40. Kloppenborg RP, Nederkoorn PJ, Geerlings MI, Berg E van den.
Presence and progression of white matter hyperintensities and cognition:
A meta-analysis. Neurology. 2014;82(23):2127--38.
